\textcolor{blue}{A gradient-based optimization method for motion planning}$^{1}$

### 2016-01-27 paper.tex
Finish Section 3.6

As illustrated in Figure~\ref{optim_grad}, let 
us denote by $\tcolli$ the abscissa of the first collision detected on path 
$\xx_{i+1}$, which previous iteration $\xx_i$ was collision-free. Thus in 
configuration $\xx_{i+1}(\tcolli)$ a collision has been 
detected. Two cases are possible:
\begin{enumerate}
\item the collision occurred between two bodies of the robot: $\body_1$ and $
\body_2$, or
\item the collision occurred between a body of the robot $\body_1$ and the 
environment.
\end{enumerate}
In the rest of this section, the first case will only be considered. Reasoning 
about the second case is similar, except that the 
constraint is on the position of $\body_1$ with respect to the environment.

@@

The principle of the method is to compute a \textcolor{blue}{linear} constraint on the initialized on the 
free-configuration $\xx_{i}(\tcolli)$ to avoid the encountered collision $\xx_{i+1}
(\tcolli)$. To handle this, we introduce a one-dimensional constraint based on the 
orthogonal direction of the encountered collision.

In the collision-configuration $\xx_{i+1}(\tcolli)$, let $\po_c\in \real^3$ be a 
contact point expressed in the global frame 
(Figure~\ref{contact-points} left).
\textcolor{red}{We denote by
  \begin{itemize}
  \item $\po_1^{loc}$ (resp. $\po_2^{loc}$) the coordinate vector of $\po_c$ in the local frame of  $\body_1$ (resp. $\body_2$),
  \item $M_1(\conf) \in SE(3)$ (resp. $M_2(\conf) \in SE(3)$) the rigid-body transformation representing the position of $\body_1$ (resp. $\body_2$) in the global frame in configuration $\conf$,
  \item $M_2^1 (\conf) = M_1(\conf)^{-1} M_2(\conf)$ the position of $\body_2$ local frame in $\body_1$ local frame in configuration $\conf$,
  \item $\po_1(\conf)$ (resp. $\po_2(\conf)$) the points moving with $\body_1$ (resp. $\body_2$) of local coordinate $\po_1^{loc}$ (resp. $\po_2^{loc}$) in $\body_1$ (resp. $\body_2$) local frame.
  %% \item $\rotation_1(\conf) \in SO(3)$ (resp. $\rotation_2(\conf) \in SO(3)$) the rotation associated with $M_1(\conf)$ (resp. $M_2(\conf)$).
  \end{itemize}
We define $\U$ as the coordinate vector of the unit vector linking points $\po_1$ and $\po_2$ in configuration $\xx_{i}(\tcolli)$, expressed in local frame of $\body_1$ (Figure~\ref{contact-points} right):
$$
\U = \frac{M_2^1 (\xx_{i}(\tcolli))\po_2^{loc} - P_1^{loc}}{\|M_2^1 (\xx_{i}(\tcolli))\po_2^{loc} - P_1^{loc}\|}
$$
Note that $\U$ is well defined since configuration $\xx_{i}(\tcolli)$ is collision-free.
}
\textcolor{blue}{
  Let $g$ be the real valued function mapping the projection of vector $\overrightarrow{\po_1\,\po_2}$ on $\mathbf{u}$ to a configuration $\conf$ . For any $\conf \in \CS$:
\begin{equation}\label{eq:g}
g (\conf) = \left(M_2^1 (\conf)\po_2^{loc} - P_1^{loc} | \U\right)
\end{equation}
}
%which is calculated from the last collision-free path at abscissa $\tcolli$.

@@

\subsubsection{Note about assumption~(\ref{eq:assumption-1})}%
\noindent
\textcolor{red}{As in the previous section, we assume that $\xx_i$ is collision-free and that $\xx_{i+1}$ is in collision at abscissa $\tcolli$.
According to~(\ref{eq:f}), the evaluation of the constraint function $f$ along the iteration line $\xx_i + t\alpha_i \p_i,\ t\in [0,1]$ going from $\xx_i$ to $\xx_{i+1}$ can be written as follows:
\begin{equation}\label{eq:f-along-pi}
f (\xx_i + t\alpha_i \p_i) = g \left((\xx_i + t\alpha_i \p_i)(\tcolli)\right).
\end{equation}
The argument of function $g$ above is a trajectory in the robot configuration space that we denote by $\traj$:
\begin{equation}\label{eq:trajectory}
\traj (t) = (\xx_{i} + t\, \alpha_i \p_{i}) (\tcolli), \ t \in [0,1]
\end{equation}
The trajectory $\traj(t)$ is defined by taking the constant abscissa $\tcolli$ and by 
moving from 
path $\xx_i$ along step $\p_i$. Note that configuration $\xx_{i+1}(\tcolli)$ is reached when $t=1$.
Substituting~(\ref{eq:trajectory}) into~(\ref{eq:f-along-pi}) and differentiating with respect  to $t$ yields
\begin{align}
f (\xx_i + t\alpha_i \p_i) &= g (\traj (t)) \\
\label{eq:df/dx}
\alpha_i \frac{\partial f}{\partial \xx} (\xx_i + t\alpha_i \p_i)\p_i &=
\frac{d}{dt}g (\traj (t))
\end{align}
From the definition~(\ref{eq:g}) of $g$, the right hand side of~(\ref{eq:df/dx}) represents the velocity of point $P_2$ in reference frame $\body_1$ projected on vector $\U$.
}

@@

\textcolor{red}{
When the new constraint is not linearly independent from the set of previous constraints, the algorithm enters an additional loop in order to find a new constraint that is linearly independent. This loop is performed by Algorithm~\ref{algo:solveRedundant} and illustrated in Figure~\ref{fig:convergence-diagram}. The loop keeps looking for paths along line segment $[\xx_{i},\xx_{i+1}]$ by dichotomy. A pair containing the latest free path and the latest path in collision, denoted by $(\xx_{Free}, \xx_{Coll})$ is stored along the loop. New iterations are chosen in the middle of this pair. 
\begin{itemize}
\item If the new path is collision-free, it replaces $\xx_{Free}$ in the pair.
\item If the new path is in collision, it replaces $\xx_{Coll}$ in the pair.
\end{itemize}
In both cases, a new constraint is built following the method described in Section~\ref{subsubsec:new-constraint}. Two cases are then possible:
\begin{enumerate}
\item at some point in the loop the new constraint is linearly independent from the previous constraints. The new constraint is added to $\Jf_{i}$ to give rise to $\Jf_{i+1}$ and the loop is interrupted, or
\item each new constraint is linearly dependent from the previous constraints and the loop never ends.
\end{enumerate}
In the second case, the iterations of the loop converge to a path that we denote by $\bar{\xx}$. $\xx_{Free}$ and $\xx_{Coll}$ also both converge to $\bar{\xx}$. Note that $\bar{\xx}$ lies at the boundary between free paths and paths in collision along the search line.
}

>>>>

As illustrated in Figure~\ref{optim_grad}, let 
us denote by $\tcolli$ the abscissa of the first collision detected on path 
$\xx_{i+1}$, which previous iteration $\xx_i$ was collision-free. Thus in 
configuration $\xx_{i+1}(\tcolli)$ a collision has been 
detected. Two cases are possible:
\begin{enumerate}
\item the collision occurred between two bodies of the robot: $\body_1$ and $
\body_2$, or
\item the collision occurred between a body of the robot $\body_1$ and the 
environment.
\end{enumerate}
In the rest of this section, the first case only will be considered. Reasoning 
about the second case is similar, except that the 
constraint is on the position of $\body_1$ with respect to the environment.

@@

The principle of the method is to compute a \textcolor{blue}{linear} constraint initialized on the 
free configuration $\xx_{i}(\tcolli)$ to avoid the encountered collision $\xx_{i+1}
(\tcolli)$. To handle this, we introduce a one-dimensional constraint based on the 
orthogonal direction of the encountered collision.

In the collision-configuration $\xx_{i+1}(\tcolli)$, let $\po_c\in \real^3$ be a 
contact point expressed in the global frame 
(Figure~\ref{contact-points} left).
\textcolor{red}{We denote by
  \begin{itemize}
  \item $\po_1^{loc}$ (resp. $\po_2^{loc}$) the coordinate vector of $\po_c$ in the local frame of  $\body_1$ (resp. $\body_2$),
  \item $M_1(\conf) \in SE(3)$ (resp. $M_2(\conf) \in SE(3)$) the rigid-body transformation representing the position of $\body_1$ (resp. $\body_2$) in the global frame, in configuration $\conf$,
  \item $M_2^1 (\conf) = M_1(\conf)^{-1} M_2(\conf)$ the position of $\body_2$ local frame in $\body_1$ local frame in configuration $\conf$,
  \item $\po_1(\conf)$ (resp. $\po_2(\conf)$) the points moving with $\body_1$ (resp. $\body_2$) of local coordinate $\po_1^{loc}$ (resp. $\po_2^{loc}$) in $\body_1$ (resp. $\body_2$) local frame.
  %% \item $\rotation_1(\conf) \in SO(3)$ (resp. $\rotation_2(\conf) \in SO(3)$) the rotation associated with $M_1(\conf)$ (resp. $M_2(\conf)$).
  \end{itemize}
We define $\U$ as the coordinate vector of the unit vector linking points $\po_1$ and $\po_2$ in configuration $\xx_{i}(\tcolli)$, expressed in local frame of $\body_1$ (Figure~\ref{contact-points} right):
$$
\U = \frac{M_2^1 (\xx_{i}(\tcolli))\po_2^{loc} - P_1^{loc}}{\|M_2^1 (\xx_{i}(\tcolli))\po_2^{loc} - P_1^{loc}\|}
$$
Note that $\U$ is well defined since configuration $\xx_{i}(\tcolli)$ is collision-free.
}
\textcolor{blue}{
  Let $g$ be the real valued function mapping the projection of vector $\overrightarrow{\po_1\,\po_2}$ on $\mathbf{u}$ to a configuration $\conf$ . For any $\conf \in \CS$:
\begin{equation}\label{eq:g}
g (\conf) = \left(M_2^1 (\conf)\po_2^{loc} - P_1^{loc} | \U\right)
\end{equation}
}
%which is calculated from the last collision-free path at abscissa $\tcolli$.

@@

\subsubsection{Note about assumption~(\ref{eq:assumption-1})}%
\noindent
\textcolor{red}{As in the previous section, we assume that $\xx_i$ is collision-free and that $\xx_{i+1}$ is in collision at abscissa $\tcolli$.
According to~(\ref{eq:f}), the evaluation of the constraint function $f$ along the iteration line $\xx_i + t\alpha_i \p_i,\ t\in [0,1]$ going from $\xx_i$ to $\xx_{i+1}$ can be written as follows:
\begin{equation}\label{eq:f-along-pi}
f (\xx_i + t\alpha_i \p_i) = g \left((\xx_i + t\alpha_i \p_i)(\tcolli)\right).
\end{equation}
The argument of function $g$ above is a trajectory in the robot configuration space that we denote by $\traj$:
\begin{equation}\label{eq:trajectory}
\traj (t) = (\xx_{i} + t\, \alpha_i \p_{i}) (\tcolli), \ t \in [0,1]
\end{equation}
The trajectory $\traj(t)$ is defined by taking the constant abscissa $\tcolli$ and by 
moving from 
path $\xx_i$ along step $\p_i$. Note that configuration $\xx_{i+1}(\tcolli)$ is reached when $t=1$.
Substituting~(\ref{eq:trajectory}) into~(\ref{eq:f-along-pi}) and differentiating with respect  to $t$ yields
\begin{align}
f (\xx_i + t\alpha_i \p_i) &= g (\traj (t)) \\
\label{eq:df/dx}
\alpha_i \frac{\partial f}{\partial \xx} (\xx_i + t\alpha_i \p_i)\p_i &=
\frac{d}{dt}g (\traj (t))
\end{align}
\begin{property}\label{prop:geometric-reasoning}
From the definition~(\ref{eq:g}) of $g$, the right hand side of~(\ref{eq:df/dx}) represents the velocity of point $P_2$ in reference frame $\body_1$ projected on vector $\U$.
\end{property}
}

@@

\textcolor{red}{
When the new constraint is not linearly independent from the set of previous constraints, the algorithm enters an additional loop in order to find a new constraint that is linearly independent. This loop is performed by Algorithm~\ref{algo:solveRedundant} and illustrated in Figure~\ref{fig:convergence-diagram}. The loop keeps looking for paths along line segment $[\xx_{i},\xx_{i+1}]$ by dichotomy. A pair containing the latest free path and the latest path in collision, denoted by $(\xx_{Free}, \xx_{Coll})$ is stored along the loop. New iterations are chosen in the middle of this pair. 
\begin{itemize}
\item If the new path is collision-free, it replaces $\xx_{Free}$ in the pair.
\item If the new path is in collision, it replaces $\xx_{Coll}$ in the pair.
\end{itemize}
In both cases, a new constraint is built following the method described in Section~\ref{subsubsec:new-constraint}. Two cases are then possible:
\begin{enumerate}
\item at some point in the loop the new constraint is linearly independent from the previous constraints. The new constraint is added to $\Jf_{i}$ to give rise to $\Jf_{i+1}$ and the loop is interrupted, or
\item each new constraint is linearly dependent from the previous constraints and the loop never ends.
\end{enumerate}
In the second case, the iterations of the loop converge to a path that we denote by $\bar{\xx}$. $\xx_{Free}$ and $\xx_{Coll}$ also both converge to $\bar{\xx}$. Note that $\bar{\xx}$ lies at the boundary between free paths and paths in collision along the search line. $\bar{\xx}$ necessarily lies at the boundary between free paths and paths in collision. Let us denote by $\bar{\kappa}$ the abscissa along $\bar{\xx}$ where $\body_1$ and $\body_2$ come to contact, and let us denote by $\bar{\po}$ the contact point.
}

\textcolor{red}{
At each iteration, the new linear constraint
$$
L_{j+1} = \frac{\partial f}{\partial\xx} (\xx_{Free})
$$
is tested. As iterations $\xx_{Free}$ and $\xx_{Coll}$ tend toward $\bar{\xx}$, it can be established by a geometric reasoning analogous to Property~\ref{prop:geometric-reasoning}, that
$L_{j+1} \p_i$ tends to the norm of the velocity of point $\bar{\po}$ belonging to $\body_2$ in the frame of $\body_1$. The following property summarizes the result of this section.
}

\textcolor{red}{
\begin{property}\label{prop:convergence}
As long as along iteration $\p_i$, the trajectory $(\xx_i+t\alpha_i\p_i)(\bar{\kappa})$ does not enter in collision with contact point velocity equal to 0 in the frame of $\body_1$, Algorithm~\ref{algo:solveRedundant} converges in a finite number of steps.
\end{property}
The above property means that the gradient-based path length optimization algorithm described in this paper converges in all cases, except for ill-defined problems.
}
